---
title: "Testing of Statistical Hypothesis"
author: "Adejumo Ridwan Suleiman"
date: "`r Sys.Date()`"
output: beamer_presentation
---

## What is a Statistical Hpothesis

* This is a formal statement about the nature of a population.
* Example; 
  - Are all Africans rich?
  - Do all footballers make above $100k/week.
  - Do Ronaldo and Messi have the same goal ratio.
  
## Hypothesis Testing

* This is the use of statistical tests to make an inference about a population.
* The method of inference used depends on the nature of the data and reason for analysis.

## Steps in conducting a Statistical Hypothesis Test

* Stating the Null and Alternative Hypothesis
* Stating the Level of Significance
* Test Statistics to be used
* Critical Region
* Decision Rule

## Null Hypothesis

* Know as the Hypothesis of no difference.
* This Hypothesis states that the relationship the researcher is trying to investigate does not exist.
* OUr hope is always to prove the **Null hypothesis** wrong.
* The **Null Hypothesis** is represented mathematically as **$H_0$**

## Alternative Hypothesis

* Opposite of the **Null Hypothesis**
* This is usually the hypothesis of the researcher.
* Represented as **$H_A$**

## Case Study

* Let's say the mean price of diamonds is less than $10000
* In this case the **Null Hypothesis** can be stated as;  
*"The mean price of diamonds is $10000"*
* The **Alternative hypothesis** can be stated as ;  
*"The mean price of diamonds is less than $1000"*
* The parameter of interest in this example is the **Mean**.

## Type 1 and Type 2 error

* Situation when you reject the **Null Hypothesis** when it is actually true is know as **Type 1** error.
* Situation when you fail to reject the **Alternative Hypothesis** when it is actually false is know as **Type 2** error.

## Level of Significance ($\alpha$)

* Situation where an experiment occurred beyond chance, then it is said to **Statistically Significant**.
* If we are to perform an experiment a number of times and the result of that experiment does not occurs by chance the experiment is declared **Statistically Significant**
* The level of significance denoted by $\alpha$ is the probability of rejecting the null hypothesis when it is actually true.
* The level of significance is mostly used as **0.05**.
* We can say that this is the probability of committing a **Type 1** error.

## **p**-Value

* This is the probability of having the result of an experiment by chance when reproduced many times.
* For example a p-value of 0.32 means that when the experiment is repeated a number of times, there is a 32% chance that the result of your experiment actually occurred by chance.
* The lower the p-value the higher the statistical significance of the test.

## Correlation Test

* This statistics test if two continuous variables are related to each other.
* The value ranges from -1 to +1 and the higher the value the stronger the relationship.
* Let's say from the diamonds data example we want to see if there is a relationship between diamonds carat and price.
```{r, echo=F, warning=F,message=F}
library(tidyverse)
cor.test(diamonds$price, diamonds$carat)
```

## One-Sample t-test

* Use to see if the mean of a sample from a population is equal to a specific value.
* Let's assume someone argues with me that the mean price of diamonds is $6000 and I am not capable to get the mean price of all diamonds from the population.
* Instead I take a random sample of 100 diamonds from the population and perform a t-test and compare it with the mean which we don't know if it is true or not.
* We can state the **Null hypothesis** as:  
*The mean price of diamonds is $6000*
* We can state the **Alternative hypothesis** as:    
*The mean price of diamonds is not $6000*

## One-Sample t-test(continuation)


```{r, echo=F, warning=F,message=F}
set.seed(1999)
t.test(diamonds[sample(nrow(diamonds), 100), ]$price, mu = 6000)
```

## One-Sample t.test(continuation)

### What if I am to use the compare the actually mean price of diamonds or a value closer to that. Let's try 4000(the actual mean is 3900)
```{r, echo=F, warning=F,message=F}
set.seed(1999)
t.test(diamonds[sample(nrow(diamonds), 100), ]$price, mu = 4000)
```
* Can you see the difference?

## Two-Sample t-test 

* This is like the one-sample t-test but is used to compare the means of two samples.
* This time around I have 100 samples of diamonds from two different store.
* I want to know if the mean price from these store are the same.
* We can state our **Null Hypothesis** as:  
*The average price of diamonds from Store A is the same as that of Store B*
* The **Alternative Hypothesis** takes the form:  
*The average price of diamonds from Store A is not the same as that of Store B*

## Two-Sample t-test(continuation)

```{r, echo=F, warning=F,message=F}
set.seed(1999)
Store_A = diamonds[sample(nrow(diamonds), 100), ]$price
Store_B = diamonds[sample(nrow(diamonds), 100), ]$price

t.test(Store_A, Store_B)
```

## Analysis of Variance (ANOVA)

* Now I have four samples from four different Store, what do I do?
* The **ANOVA** is used when you want to compare the mean of more than two samples.
* We can state our **Null Hypothesis** as:  
*The mean price of diamonds is the same across all Stores*
* The **Alternative Hypothesis** takes the form:  
*The average price of diamonds is not the same across all Stores*

```{r, echo=F, warning=F,message=F}
set.seed(1999)
Store_A = data.frame(Price =diamonds[sample(nrow(diamonds), 100), ]$price,
                     Store = rep(1,100))

Store_B = data.frame(Price =diamonds[sample(nrow(diamonds), 100), ]$price,
                     Store = rep(2,100))
Store_C = data.frame(Price =diamonds[sample(nrow(diamonds), 100), ]$price,
                     Store = rep(3,100))
Store_D = data.frame(Price =diamonds[sample(nrow(diamonds), 100), ]$price,
                     Store = rep(4,100))

df <- rbind(Store_A,Store_B,Store_C,Store_D)

aov_results <- aov(df$Price ~ df$Store)
summary(aov_results)
```















